---
subtitle: "Pattern Mining and Social Network Analysis"
title: "Homework 3"
author: "BOUYSSOU Gatien , de POURTALES Caroline, LAMBA Ankit"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 6
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=6, fig.height=5)
library(reticulate)
#use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = T)
knitr::knit_engines$set(python.reticulate =  TRUE)
#py_install("matplotlib")
#py_install("scikit-learn")

#install_LUCS_KDD_CMAR(force = TRUE,source = "https://cgi.csc.liv.ac.uk/~frans/KDD/Software/CMAR/cmar.tgz")
#install_LUCS_KDD_CPAR(force = TRUE,source = "https://cgi.csc.liv.ac.uk/~frans/KDD/Software/FOIL_PRM_CPAR/foilPrmCpar.tgz")

```

```{r rpackages,eval=TRUE,echo=FALSE}
#install.packages("fpc")
#install.packages("arules")
#install.packages("arulesCBA")
#install.packages("mlbench")
library(mlbench)
library(magrittr)
library(knitr)
library(arulesCBA)
library(rmarkdown)
library(arules)
library(arulesViz)
library(cluster)
library(caret)
```


\clearpage

# Parameters in association rules

There are parameters controlling the number of rules to be generated.

## Support

Support is an indication of how frequently the itemset appears in the dataset.

$$Support(A→B) = \frac{\text{Number of transaction with both A and B}}{\text{Total Number of transaction}} = P(A \cap B)$$

## Confidence

Confidence is an indication of how often the rule has been found to be true.\newline
This says how likely B is induced by A.

$$Confidence (A→B)= \frac{\text{Number of transaction with both A and B}}{\text{Total Number of transaction with A}} = \frac{P(A \cap B)}{P(A)}$$

## Lift

Lift is the factor by which, the co-occurence of A and B exceeds the expected probability of A and B co-occuring, had they been independent. So, higher the lift, higher the chance of A and B occurring together.

$$Lift(A→B) = \frac{P(A \cap B)}{P(A) × P(B)}$$

## Leverage

The leverage compares the frequency of A and B appearing together and the frequency that would be expected if A and B were independent. 

$$Levarage(A→B)= P(A \cap B) − P(A)×P(B)$$

Therefore, if A and B independent : 

$Levarage(A→B) = 0$

## Conviction

Conviction compares the probability that A appears without B if they were dependent with the actual frequency of the appearance of A without B. In that respect it is similar to lift, however, it contrast to lift it is a directed measure. Furthermore, conviction is monotone in confidence and lift.

$$Conviction(A → B)=\frac{P(A) × P(\bar{B})}{P(A \cap \bar{B})}$$

or

$$Conviction(A → B)=\frac{1 - P(B)}{1- \frac{P(A \cap B)}{P(A)}}$$

\clearpage

# APRIORI algorithm

## Definition

APRIORI searches for frequent itemset browsing the lattice of itemsets in breadth. \newline
The database is scanned at each level of lattice. Additionally, APRIORI uses a pruning technique based on the properties of the itemsets, which are: If an itemset is frequent, all its sub-sets are frequent and not need to be considered.

## Example on Groceries data on R

The Groceries data set contains 30 days of real-world point-of-sale transaction data from a typical local grocery outlet. The data set contains 9835 transactions and the items are aggregated to 169 categories.

We can see the class of the dataset is : 
```{r}
data("Groceries")
class(Groceries)
```

Looking at some examples of transaction :

```{r}
inspect(head(Groceries,3))
```

We can find the 15 most common variables.
```{r, eval=TRUE, echo=FALSE, fig.height = 6, fig.width = 8, fig.align = "center"}
itemFrequencyPlot(Groceries, topN=15, type="absolute", main="Item Frequency")
```

Let's apply APRIORI algorithm on the dataset : 
```{r, eval=TRUE, echo=FALSE}
grocery_rules <- apriori(Groceries, parameter = list(support = 0.005, confidence = 0.2))
```


\clearpage 

We have a set of associations rules. 
```{r, eval=TRUE, echo=FALSE}
grocery_rules
```

If we look at the 3 rules with highest confidence, we have these rules : 
```{r, eval=TRUE, echo=FALSE}
inspect(head(sort(grocery_rules, by = "confidence", decreasing=TRUE), 3))
```
However "whole milk" if the most frequent item in the data set and this frequence plays a role in confidence.
So we can look at the 3 rules with highest lift (A and B occuring together), and we have these rules : 

```{r, eval=TRUE, echo=FALSE}
inspect(head(sort(grocery_rules, by = "lift", decreasing=TRUE), 3))
```

We can also look at the items which induce "soda". Then we can sort them by confidence and look at the first 3 (so the 3 rules with highest confidence).

```{r, eval=TRUE, echo=FALSE}
inspect(head(sort(subset(grocery_rules, subset = rhs %in% "soda"), by = "confidence", decreasing=TRUE), 3))
```

Looking at the confidence, we see that for a third of the people buying : 
- bottled water and fruit/vegetable juice or 
- sausage and shopping bags or 
- yogurt and bottled water

it also induces buying soda.

\clearpage

# Using Frequent itemsets to find rules

## Concept

\textbf{Closed Frequent itemsets :}

An itemset X is a closed frequent itemset in set S if X is both closed and frequent in S.

\textbf{Eclat algorithm :}

It mines frequent itemsets \newline
This algorithm uses simple intersection operations for equivalence class clustering along with bottom-up lattice traversal.
Then looking at the most frequent itemsets, we can find rules between the items inside these itemsets.

## Example on Adult data on R

The Adult data set from R contains 48842 observations on the 15 variables (age, workclass, ...).

```{r, eval=TRUE, echo=FALSE}
data("Adult")
class(Adult)
```

We can look at the first transaction to see what are the items in a transaction. 
```{r, eval=TRUE, echo=FALSE}
inspect(head(Adult,1))
```

We can find the 15 most common variables.
```{r, eval=TRUE, echo=FALSE, fig.height = 6, fig.width = 8, fig.align = "center"}
itemFrequencyPlot(Adult, topN=15, type="absolute", main="Item Frequency")
```

We apply ECLAT algorithm on the data set. \newline
This returns the most frequent itemsets along with their support.\newline
Let's look at 3 of these itemsets : 
```{r, eval=TRUE, echo=FALSE, fig.height = 6, fig.width = 8}
frequentItemsAdult <- eclat (Adult, parameter = list(supp = 0.01, maxlen = 100))
inspect(head(frequentItemsAdult,3))
```

We can also find the rules from the most frequent itemsets. \newline
We use the ruleInduction function from R. We can set the method with the argument "method".

If in control method = "APRIORI" is used, a very simple rule induction method is used. All rules are mined from the transactions data set using APRIORI with the minimal support found in itemsets. And in a second step all rules which do not stem from one of the itemsets are removed. This procedure will be in many cases very slow.

```{r, fig.height = 6, fig.width = 8}
rulesAdult <- ruleInduction(frequentItemsAdult, confidence = 0.85, control = list(method = "apriori"))
inspect(head(sort(rulesAdult, by = "lift", decreasing=TRUE), 3))
```

If in control method = "ptree" is used, the transactions are counted into a prefix tree and then the rules are selectively generated using the counts in the tree. This is usually faster than the above approach.


We can also find the rules with a specific given result. \newline
For example, let's answer the question : \newline
How to be rich ?

```{r, fig.height = 6, fig.width = 8}
frequentItemsAdultGain <- eclat (Adult, parameter = list(supp = 0.01, maxlen = 200))
rulesAdult <- ruleInduction(frequentItemsAdultGain, confidence = 0.15, control = list(method = "apriori"))
rulesAdult <- subset(rulesAdult, rhs %pin% "capital-gain=High")
```


We take the 3 best rules according to lift.  

```{r}
rulesAdult
inspect(head(sort(rulesAdult, by = "lift", decreasing=TRUE), 3))
```

We see a pattern for people with a high capital gain : they have often a large income, work over-time and have no capital loss. 


TO DO : difference between ECLAT and APRIORI.

\clearpage

## Example on mushroom data on python with scikit-learn

This database contains a lot of mushrooms with a set of characteristics. Each mushroom is classified either as edible or poisonous. The database has been found in kaggle and is available here : https://www.kaggle.com/uciml/mushroom-classification.

```{python}
import pandas as pd
import urllib
import matplotlib.pyplot as plt
import numpy as np
import os
from mlxtend.frequent_patterns import apriori, association_rules, fpgrowth
from mlxtend.preprocessing import TransactionEncoder
```


```{python}
mush_data = pd.read_csv('./dataPython/mushrooms.csv')
```

First, we want to have an overview of the data.

```{python}
mush_data.head()
```
As we can see, each column contains values that are single characters. Their meaning is given by the file values_name.txt.

```{python}
len(mush_data)
```

Now, we want to know the data repartition for each columns.

```{python}
for column in mush_data.columns:
    print("\n" + column)
    print(mush_data[column].value_counts())
```

As you can see the there is almost as much poisonous as edible mushrooms. Moreover, the dataset contains some unknown values in the column stalk-root. We are going to discard those rows to keep lines that are complete.

```{python}
mush_data = mush_data[mush_data["stalk-root"] != '?']
```


```{python}
len(mush_data)
```

```{python}
mush_data['class'].value_counts()
```

Even without the discarded lines the dataset still have plenty of data and the class label is almost balanced. Before feeding the APRIORI algorithm with our data, we need to use the TransactionEncoder provided by mlxtend. This class transforms our data into a matrix where :

- each possible value for each feature will become a column
- for each mushroom and each column we assign a boolean that correspond to weither or not the feature is contained by the mushroom.

For example, such a dataset :

\begin{table}[]
\begin{tabular}{ll}
Columns : & odor    \\
0         & pungent \\
1         & almond  \\
2         & anise   \\
3         & pungent \\
4         & none
\end{tabular}
\end{table}

Will be changed into this matrix :

\begin{table}[]
\begin{tabular}{lllll}
Columns : & pungent & almond & anise & none  \\
0         & True    & False  & False & False \\
1         & False   & True   & False & False \\
2         & False   & False  & True  & False \\
3         & True    & False  & False & False \\
4         & False   & False  & False & True
\end{tabular}
\end{table}

The mushroom dataset contains characters values. In order to have columns that are bit more intelligible we will replace the character values by their full name.

This function above split for one feature the one character values from the full name values. It returns two arrays with each type of values.

```{python}
def sepCurrentValFromNewVal(stringNewOldValues):
    currentVal = []
    newVal = []
    arrayNOvalues = stringNewOldValues.split(",")
    for NOValues in arrayNOvalues:
        NO = NOValues.split('=')
        currentVal.append(NO[1])
        newVal.append(NO[0])
    return [currentVal, newVal]
```

This function goes through all the features and maps the feature's names with the character and full name values. Therefore it changes this line :

cap-shape: bell=b,conical=c,convex=x,flat=f, knobbed=k,sunken=s

into that dictionary :

{'cap-shape' : [['b', 'c', 'x', 'f', 'k', 's'],
 ['bell', 'conical', 'convex', 'flat', 'knobbed', 'sunken']]}

Then, we will replace the values in the first array by the values of the second one for a given column.

```{python}
mapColumnValName = {}
with open("./dataPython/values_name.txt", "r") as file:
    line = file.readline()
    while line != "":
        line = line.replace(' ', "").replace("\n", "")
        valuesNamesForColumn = line.split(':')
        mapColumnValName[valuesNamesForColumn[0]] = sepCurrentValFromNewVal(valuesNamesForColumn[1])
        line = file.readline()
mapColumnValName['cap-shape']
```


```{python}
for column, ONValues in mapColumnValName.items():
    mush_data[column] = mush_data[column].replace(ONValues[0], ONValues[1])
mush_data.head()
```

```{python}
train_data = mush_data.values
```

```{python}
te = TransactionEncoder()
te_ary = te.fit(train_data).transform(train_data)
df = pd.DataFrame(te_ary, columns=te.columns_)
len(df.head())
```

```{python}
frequent_itemsets = apriori(df, min_support=0.6, use_colnames=True)
frequent_itemsets['length'] = frequent_itemsets['itemsets'].apply(lambda x: len(x))
frequent_itemsets
```

Thanks to the APRIORI algorithm it is possible to associate some feature together.

```{python}
frequent_itemsets[frequent_itemsets['itemsets'].astype(str).str.contains("edible")]
```

```{python}
frequent_itemsets[frequent_itemsets['itemsets'].astype(str).str.contains("poisonous")]
```

With the APRIORI algorithm, we can see some associations with the \textit{edible} feature with a support around 0.6. Also, it seems that the APRIORI haven't found any associations with the \textit{poisonous} feature with a support above 0.6.

The result given by APRIORI will be used by the association_rules function given by mixtend.

```{python}
assos_rule = association_rules(frequent_itemsets, metric="confidence", min_threshold=0.9)
```

```{python}
assos_rule[assos_rule['antecedents'].astype(str).str.contains("edible")]
```
Here we are listing all the rules that are implied by edible. Of course, we need to know the rules where edible is implied (ie the rules where edible is contained by the consequents column). But before searching for those rules, we are going to try out another algorithm, named fpgrowth, to see if we can obtain different results.

```{python}
frequent_itemsets = fpgrowth(df, min_support=0.6, use_colnames=True)
```

```{python}
frequent_itemsets[frequent_itemsets['itemsets'].astype(str).str.contains("edible")]
```

The results above obtained by fpgrowth look similar to the results obtained by the APRIORI algorithm. Therefore, now we can look for the rules that implies edible.


```{python}
assos_rule = association_rules(frequent_itemsets, metric="confidence", min_threshold=0.7)
```

```{python}
assos_rule[assos_rule['consequents'].astype(str).str.contains("edible")]
```

```{python}
assos_rule = association_rules(frequent_itemsets, metric="confidence", min_threshold=0.6)
```


```{python}
assos_rule[assos_rule['consequents'].astype(str).str.contains("edible")]
```

As we can see above if the threshold is above 0.6 the association_rules function does not find any rules where edible is implied. The confidence and the consequent support of the rules lies around 60%. Therefore, they cannot be considered as reliable.

\clearpage

# Clustering with APRIORI algorithm

## Concept

We can find a dissimilarity between transactions so we can compare the data. Then this dissimilarity is used as distance measure in clustering.

So a direct approach to cluster itemsets is to define a distance metric between two itemsets $X_i$ and $X_j$.

## Affinity dissimilarity 

A good choice is the Affinity defined as :

$$A(X_i, X_j)=\frac{Support(X_i,X_j)}{P(X_i) + P(X_j) - Support(X_i,X_j)} = \frac{P(X_i \cap X_j)}{P(X_i \cup X_j)}$$

Here this means that affinity is the Jaccard similarity between items. \newline
The Jaccard distance defined as :

$$J(X_i, X_j) = \frac{|X_i \cap X_j|}{|X_i \cup X_j|}$$

The distance simply is the number of items that $X_i$ and $X_j$ have in common divided by the number of unique items in both sets.

## Example on tennis data on R

We use a dataset from the Wimbledon tennis tournament for Women in 2013. We will predict the result for player 1 (win=1 or loose=0) based on : the number of aces won by each player, and, the number of unforced errors commited by both players. The data set is a subset of a data set from https://archive.ics.uci.edu/ml/datasets/Tennis+Major+Tournament+Match+Statistics. \newline

```{r}
id <- "1GNbIhjdhuwPOBr0Qz82JMkdjUVBuSoZd"
tennis.data <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download",id), header = T)
tennis.data <- tennis.data[,3:7]
tennis.data <- data.frame(tennis.data)
head(tennis.data)
```

We can transform the tennis data set into a transaction data set. 
```{r, eval=TRUE, echo=FALSE, fig.height = 5, fig.width = 10, fig.align = "center"}
tennisTransactions <- tennis.data

tennisTransactions[["ACE.1"]] <- ordered(cut(tennisTransactions[[ "ACE.1"]],
  c(-Inf,0, median(tennisTransactions[[ "ACE.1"]][tennisTransactions[[ "ACE.1"]]>0]),
  Inf)), labels = c("None", "Low", "High"))

tennisTransactions[["ACE.2"]] <- ordered(cut(tennisTransactions[[ "ACE.2"]],
  c(-Inf,0, median(tennisTransactions[[ "ACE.2"]][tennisTransactions[[ "ACE.2"]]>0]),
  Inf)), labels = c("None", "Low", "High"))

tennisTransactions[["UFE.2"]] <- ordered(cut(tennisTransactions[[ "UFE.2"]],
  c(-Inf,0, median(tennisTransactions[[ "UFE.2"]][tennisTransactions[[ "UFE.2"]]>0]),
  Inf)), labels = c("Low", "High"))

tennisTransactions[["UFE.1"]] <- ordered(cut(tennisTransactions[[ "UFE.1"]],
  c(-Inf,0, median(tennisTransactions[[ "UFE.1"]][tennisTransactions[[ "UFE.1"]]>0]),
  Inf)), labels = c("Low", "High"))

tennisTransactions[["Result"]] <- as.factor(tennisTransactions[["Result"]])

tennis.transactions <- as(tennisTransactions, "transactions")
```


We can look at the 3 first transactions.
```{r}
inspect(head(tennis.transactions,3))
```

We can restrict the rules to the result rhs="Result=1" which means Player-1 winner.

The associations rules for Player-1 winning are :

```{r, eval=TRUE, echo=FALSE}
tennis.rules_winning <- apriori(tennis.transactions, appearance = list (default="lhs",rhs="Result=1"), parameter = list(support = 0.15, confidence = 0.3))
```

```{r, eval=TRUE, echo=FALSE}
tennis.rules_winning
inspect(head(sort(tennis.rules_winning, by = "lift", decreasing=TRUE), 5))
```

These rules look correct : either a player-1 winning make a lot of aces and few unforced errors or the player-2 make few aces.


We can also restrict the rules to the result rhs="Result=0" which means Player-1 loosing.

The associations rules for Player-1 loosing :

```{r, eval=TRUE, echo=FALSE}
tennis.rules_loosing <- apriori(tennis.transactions, appearance = list (default="lhs",rhs="Result=0"), parameter = list(support = 0.15, confidence = 0.3))
```

```{r, eval=TRUE, echo=FALSE}
tennis.rules_loosing
inspect(head(sort(tennis.rules_loosing, by = "lift", decreasing=TRUE), 3))
```

These rules look correct : either player-1 is loosing because player-2 makes a lot of aces or because he does a lot of unforced errors or player-2 makes a lot of aces.


Now let's look at all the associations rules leading to "Result". \newline
All the rules with Result as association :

```{r, eval=TRUE, echo=FALSE}
tennis.rules <- apriori(tennis.transactions, appearance = list (default="lhs",rhs=list("Result=0","Result=1")), parameter = list(support = 0.2, confidence = 0.4))
```

```{r, eval=TRUE, echo=FALSE}
tennis.rules
inspect(head(sort(tennis.rules, by = "lift", decreasing=TRUE), 5))
```

Firstly let's look at the clustering of items 

### Cluster the items

```{r fig.height=5, fig.width=6, fig.align = "center"}
d <- dissimilarity(tennis.transactions,method = "Jaccard", which = "items")
clusters_hiearchical <- hclust(d, method = "ward.D2")
plot(clusters_hiearchical)
```

We can see two clusters on the dendogramm. One contains Result=1 and the other contains Result=0. \newline
Moreover in the cluster with the branch Result=1, we can also see that the number of aces made by player-1 is high and the number of unforced errors is low. Also in this cluster player-2 made few or none aces. \newline
In the other cluster, it is the opposite. Player-1 makes few or none aces whereas player-2 makes a lot.

So it seems that this clustering manage to cluster data linked to the result together.


Now let's try to cluster the rules.

\clearpage

### Cluster the rules

```{r fig.height=5, fig.width=6, fig.align = "center"}
d <- dissimilarity(tennis.rules,method = "Jaccard", which = "associations")
clusters_hiearchical <- hclust(d, method = "ward.D2")
plot(clusters_hiearchical)
```

If we cut the dendogramm in two clusters. We can look at the first cluster. (We only print 5 items from the cluster, look at the code for the wole cluster)
```{r , eval =TRUE, fig.height=5, fig.width=6, fig.align = "center"}
assign <- cutree(clusters_hiearchical, k=2)
inspect(head(tennis.rules[assign == 1],5))
```

```{r , eval =FALSE, fig.height=5, fig.width=6, fig.align = "center"}
#Execute this to see the wholde cluster
inspect(tennis.rules[assign == 1])
```

And at the second cluster. (We only print 5 items from the cluster, look at the code for the wole cluster)
```{r , eval =TRUE, fig.height=5, fig.width=6, fig.align = "center"}
inspect(head(tennis.rules[assign == 2],5))
```

```{r , eval =FALSE, fig.height=5, fig.width=6, fig.align = "center"}
#Execute this to see the wholde cluster
inspect(tennis.rules[assign == 2])
```

This clustering regroups Player-1 winner together and Player-2 winner together.

\clearpage

# Association Rule Classification

## Classification Based on Associations : CBA Algorithm

CBA (Classification Based on Associations) Algorithm build a classifier based on association rules mined for an input dataset. 
Candidate classification association rules (CARs) are mined with the standard APRIORI algorithm. Rules are ranked by confidence, support and size. 
### Example on tennis data on R

#### Recall from Homework 1

With Random Forest, the accuracy rate was 0.6931818.\newline
With Logistic regression it was 0.7667.

#### Classification using data set and chi2 discretization

CBA can take as input a classic non-transaction dataset as tennis. We just have to choose the discretization method in parameter. \newine
chi2 discretization = TO DO

```{r}
tennisClassifier <- tennis.data

tennisClassifier[["ACE.1"]] <- as.numeric(tennisClassifier[["ACE.1"]])

tennisClassifier[["ACE.2"]] <- as.numeric(tennisClassifier[["ACE.2"]])

tennisClassifier[["UFE.2"]] <- as.numeric(tennisClassifier[["UFE.2"]])

tennisClassifier[["UFE.1"]] <- as.numeric(tennisClassifier[["UFE.1"]])

tennisClassifier[["Result"]] <- as.factor(tennisClassifier[["Result"]])

# test and train set
n = dim(tennisClassifier)[1]
n2 = n*(5/6)
set.seed(1234)
train = sample(c(1:n), replace = F)[1:n2]
tennisTest = tennisClassifier[-train, ]
tennisTrain = tennisClassifier[train, ]

#Learn a classifier using automatic default discretization
classifierData <- CBA(Result ~ ACE.1 + ACE.2 + UFE.1 + UFE.2, data = tennisTrain, disc.method = "chi2", supp = 0.05, conf=0.4)
classifierData
```

```{r}
# inspect the rule base
inspect(rules(classifierData))
```

```{r}
# make predictions
classifierData.prediction = predict(classifierData, tennisTest)
classifierData.confusion_matrix = table(classifierData.prediction, true = tennisTest$Result)
classifierData.confusion_matrix
```


The accuracy rate is :
```{r, eval=TRUE, echo=FALSE}
classifierData.accuracyrate = (classifierData.confusion_matrix[1,1] + classifierData.confusion_matrix[2,2]) / (classifierData.confusion_matrix[1,1] + classifierData.confusion_matrix[1,2] + classifierData.confusion_matrix[2,1] +classifierData.confusion_matrix[2,2])
classifierData.accuracyrate
```

The sensitivity is the percentage of true output giving Player1-winner among the population of true Player1-winner :
```{r, eval=TRUE, echo=FALSE}
classifierData.sensitivity = classifierData.confusion_matrix[2,2]/(classifierData.confusion_matrix[1,2] + classifierData.confusion_matrix[2,2])
classifierData.sensitivity
```

The specificity is the percentage of true output giving Player2-winner (= Player1-looser) among the population of true Player2-winner:
```{r, eval=TRUE, echo=FALSE}
classifierData.specificity = classifierData.confusion_matrix[1,1]/(classifierData.confusion_matrix[1,1] + classifierData.confusion_matrix[2,1])
classifierData.specificity
```

The precision is the percentage of true output giving Player1-winner among all the outputs giving Player1-winner (even if not winner) :
```{r, eval=TRUE, echo=FALSE}
classifierData.precision = classifierData.confusion_matrix[2,2]/(classifierData.confusion_matrix[2,1] + classifierData.confusion_matrix[2,2])
classifierData.precision
```

So the F_Mesure is :
```{r, eval=TRUE, echo=FALSE}
classifierData.fmesure = (2*classifierData.precision*classifierData.sensitivity)/(classifierData.sensitivity + classifierData.precision)
classifierData.fmesure
```


#### Classification using homemade transactions 

We can also use the transactions we created before. \newline
ACE.1 and ACE.2 take for value either None, or Low or High. \newline
UFE.1 and UFE.2 take for value either Low or High. \newline

```{r}
#Build the classifier from transactions
tennisTransactionsTrain <- tennis.transactions[train,]
tennisTransactionsTest <- tennis.transactions[-train,]

classifierTransactions <- CBA(Result ~ ACE.1 + ACE.2 + UFE.1 + UFE.2, data = tennisTransactionsTrain, supp = 0.05, conf = 0.4)
classifierTransactions
```

```{r}
# inspect the rule base
inspect(head(rules(classifierTransactions),4))
```

```{r}
# make predictions
classifierTransactions.prediction = predict(classifierTransactions, tennisTransactionsTest)
classifierTransactions.confusion_matrix = table(classifierTransactions.prediction, response(Result ~ ACE.1 + ACE.2 + UFE.1 + UFE.2, tennisTransactionsTest))

classifierTransactions.confusion_matrix
```

The accuracy rate is :
```{r, eval=TRUE, echo=FALSE}
classifierTransactions.accuracyrate = (classifierTransactions.confusion_matrix[1,1] + classifierTransactions.confusion_matrix[2,2]) / (classifierTransactions.confusion_matrix[1,1] + classifierTransactions.confusion_matrix[1,2] + classifierTransactions.confusion_matrix[2,1] +classifierTransactions.confusion_matrix[2,2])
classifierTransactions.accuracyrate
```
So the accuracy rate is better with our homemade transaction. It makes sense because discretization of integers is difficult to handle.


The sensitivity is the percentage of true output giving Player1-winner among the population of true Player1-winner :
```{r, eval=TRUE, echo=FALSE}
classifierTransactions.sensitivity = classifierTransactions.confusion_matrix[2,2]/(classifierTransactions.confusion_matrix[1,2] + classifierTransactions.confusion_matrix[2,2])
classifierTransactions.sensitivity
```

The specificity is the percentage of true output giving Player2-winner (= Player1-looser) among the population of true Player2-winner:
```{r, eval=TRUE, echo=FALSE}
classifierTransactions.specificity = classifierTransactions.confusion_matrix[1,1]/(classifierTransactions.confusion_matrix[1,1] + classifierTransactions.confusion_matrix[2,1])
classifierTransactions.specificity
```

The precision is the percentage of true output giving Player1-winner among all the outputs giving Player1-winner (even if not winner) :
```{r, eval=TRUE, echo=FALSE}
classifierTransactions.precision = classifierTransactions.confusion_matrix[2,2]/(classifierTransactions.confusion_matrix[2,1] + classifierTransactions.confusion_matrix[2,2])
classifierTransactions.precision
```

So the F_Mesure is :
```{r, eval=TRUE, echo=FALSE}
classifierTransactions.fmesure = (2*classifierTransactions.precision*classifierTransactions.sensitivity)/(classifierTransactions.sensitivity + classifierTransactions.precision)
classifierTransactions.fmesure
```

This classification is also as good as logistic regression. It gives very good result.

#### Classification using rules 

```{r}
#Build the classifier from rules
classifierRules <- CBA_ruleset(Result ~ . ,tennis.rules)
classifierRules
```

```{r}
# make predictions
classifierRules.prediction = predict(classifierRules, tennisTransactionsTest)
classifierRules.confusion_matrix = table(classifierRules.prediction,response(Result ~ ., tennisTransactionsTest))
classifierRules.confusion_matrix
```

## Regularized Class Association Rules for Multi-class Problems : RCAR Algorithm 

### Concept

TO DO

### Example on tenis data on R

```{r}
classifierRCAR = RCAR(Result ~ ACE.1 + ACE.2 + UFE.1 + UFE.2, data= tennisTrain, lambda = NULL, alpha = 1, disc.method = "chi2")
classifierRCAR
```

```{r}
# make predictions
classifierRCAR.prediction = predict(classifierRCAR, tennisTransactionsTest)
classifierRCAR.confusion_matrix = table(classifierRCAR.prediction, response(Result ~ ACE.1 + ACE.2 + UFE.1 + UFE.2, tennisTransactionsTest))
classifierRCAR.confusion_matrix
```

## First Order Inductive Learner : FOIL Algorithm

### Concept 

a greedy algorithm that learns rules to distinguish positive from negative examples.


Implements FOIL (Quinlan and Cameron-Jones, 1995) to learn rules and then use them as a classi- fier following Xiaoxin and Han (2003).
For each class, we find the positive and negative examples and lean the rules using FOIL. Then the rules for all classes are combined and sorted by Laplace accuracy on the training data.
Following Xiaoxin and Han (2003), we classify new examples by
1. select all the rules whose bodies are satisfied by the example;
2. from the rules select the best k rules per class (highest expected Laplace accuracy);
3. averagetheexpectedLaplaceaccuracyperclassandchoosetheclasswiththehighestaverage.

### On R

```{r}
data(Glass)
str(Glass)
response(Type ~ ., Glass)
classifierFOIL <- FOIL(Type ~ ., data= Glass, max_len = 3, min_gain = .7, best_k = 5, disc.method = "chi2")
```

## Classification Based on Multiple Class-association Rules : CMAR Algorithm

```{r}
data("iris")
# install and compile CMAR
install_LUCS_KDD_CMAR()
# build a classifier, inspect rules and make predictions
classifierCMAR <- CMAR(Species ~ ., iris, support = .2, confidence = .8, verbose = TRUE)
classifierCMAR
```

## Classification based on Predictive Association Rules : CPAR Algorithm

```{r}
classifierCPAR <- CPAR(Result ~ ACE.1 + ACE.2 + UFE.1 + UFE.2, best_k = 5, data= tennisTransactionsTrain, support = 0.1, confidence = 0.5,disc.method = "mdlp", verbose = FALSE)
classifierCPAR
```

## Classification by Aggregating Emerging Patterns : CAEP Algorithm 

No implementation in R

## A Lazy Approach to Associative Classification : L Algorithm

No implementation in R

