---
subtitle: "Pattern Mining and Social Network Analysis"
title: "Homework 3"
author: "BOUYSSOU Gatien , de POURTALES Caroline, LAMBA Ankit"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 6
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3)
library(reticulate)
#use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = T)
knitr::knit_engines$set(python.reticulate =  TRUE)
#py_install("matplotlib")
#py_install("scikit-learn")
```

```{r rpackages,eval=TRUE,echo=FALSE}
#install.packages("fpc")
#install.packages("arules")
#install.packages("arulesViz")
library(magrittr)
library(knitr)
library(rmarkdown)
library(arules)
library(arulesViz)
```

\clearpage

# Association rules Mining

## Definition and concept

TO DO

## Kinds of mining 

### Frequent pattern Itemset mining

Definition 

Kinds of Frequent Pattern Itemset mining :

Closed frequent itemset

Maximal frequent itemset

Constrained frequent itemset

Approximate frequent itemset

Near-match frequent itemset

Top-k frequent itemset

### Sequential Pattern

TO DO

### Structured Pattern 

TO DO

## Kinds of association rule mining 

### Single dimensional VS Multi dimensional association rule mining

TO DO


### Boolean VS Quantitative association rule mining

TO DO

### Correlation rule 

TO DO

## Parameters in association rules

There are three parameters controlling the number of rules to be generated.

### Support and Confidence

TO DO

Support is an indication of how frequently the itemset appears in the dataset.

$$Support = \frac{\text{Number of transaction with both A and B}}{\text{Total Number of transaction}} = P(A \cap B)$$
Confidence is an indication of how often the rule has been found to be true.

$$Confidence = \frac{\text{Number of transaction with both A and B}}{\text{Total Number of transaction with A}} = \frac{P(A \cap B)}{P(A)}$$

Lift is the factor by which, the co-occurence of A and B exceeds the expected probability of A and B co-occuring, had they been independent. So, higher the lift, higher the chance of A and B occurring together.

$$Lift = \frac{P(A \cap B)}{P(A) *P(B)}$$
\clearpage

# Apriori algorithm

## Definition 

## Example on Groceries data

```{r, eval=TRUE, echo=FALSE}
data("Groceries")
inspect(head(Groceries,5))
```
```{r, eval=FALSE, echo=FALSE}
frequentItems <- eclat (Groceries, parameter = list(supp = 0.07, maxlen = 15))
```
```{r, eval=TRUE, echo=FALSE}
inspect(frequentItems)
itemFrequencyPlot(Groceries, topN=10, type="absolute", main="Item Frequency") 
```

### On R 

```{r, eval=FALSE, echo=FALSE}
grocery_rules <- apriori(Groceries, parameter = list(support = 0.01, confidence = 0.5))
```
```{r, eval=TRUE, echo=FALSE}
inspect(head(sort(grocery_rules, by = "confidence", decreasing=TRUE), 5))
```

### On Python with scikit-learn 

```{python}
data = r.Groceries
```

# Frequent pattern-based cluster analysis

## The CLIQUE algorithm

## The ENCLUS algorithm

\clearpage

# Frequent pattern-based classification

## Classification based on Association

## Classification based on Multiple Association Rules

## Classification based on Predictive Association Rules

\clearpage

# Evaluation

Compare the algorithms






