---
subtitle: "Pattern Mining and Social Network Analysis"
title: "Homework 3"
author: "BOUYSSOU Gatien , de POURTALES Caroline, LAMBA Ankit"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 6
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3)
library(reticulate)
#use_python("/Library/Frameworks/Python.framework/Versions/3.6/bin/python3", required = T)
knitr::knit_engines$set(python.reticulate =  TRUE)
#py_install("matplotlib")
#py_install("scikit-learn")
```

```{r rpackages,eval=TRUE,echo=FALSE}
#install.packages("fpc")
library(magrittr)
library(knitr)
library(rmarkdown)
library(xlsx)
library(ggplot2)
library(ggfortify)
library(MASS)
library(dplyr)
library(ISLR)
library(readr)
library(randomForest)
library(tidyverse)
library(caret)
library(cluster)
library(factoextra)
library(fpc)
```

\clearpage

# Association rules Mining

## Definition and concept

TO DO

## Properties

TO DO

## Kinds of mining 

### Frequent pattern Itemset mining

Definition 

Kinds of Frequent Pattern Itemset mining :

Closed frequent itemset

Maximal frequent itemset

Constrained frequent itemset

Approximate frequent itemset

Near-match frequent itemset

Top-k frequent itemset

### Sequential Pattern

TO DO

### Structured Pattern 

TO DO

## Kinds of association rule mining 

### Single dimensional VS Multi dimensional association rule mining

TO DO


### Boolean VS Quantitative association rule mining

TO DO

### Correlation rule 

TO DO

## Evaluation of an association rule 

TO DO

### Support and Confidence

TO DO

### Statistical criteria

TO DO

\clearpage

# Frequent pattern-based cluster analysis

## The CLIQUE algorithm

## The ENCLUS algorithm

\clearpage

# Frequent pattern-based classification

## Classification based on Association

## Classification based on Multiple Association Rules

## Classification based on Predictive Association Rules

\clearpage

# Evaluation

Compare the algorithms






